---
title: "Project proposal"
author: "Group 6"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(ggplot2)
```

## 1. Introduction

What will be the rating of a movie based on director, genre, actor and budget?
We are trying to build a movie rating recommendation system based on certain variables in the dataset.Our selection of the predictor variables are based on intuition as we think these are the factors which influence a movie rating in a real world scenario.We would like to explore these variables further and build the linear regression model to check how accurate or far off we are in choosing these variables for movie rating prediction.


The dataset has been taken from Kaggle and consists of IMDB movie information. It is a well rounded dataset with both numerical and categorical variables. We are going to analyze each and every factors which can influence the IMDB ratings so that we can predict better results.As our entire team consists of extreme movie buffs we thought we could work on predicting the rating of a movie and see how that works! Below is a brief overview of our project plan:

1.EDA of the dataset to analyze the data before coming to any assumptions
2.Build visualizations to identify patterns, trends and outliers 
3.Find a correlation between the IMDB score and other variables in the dataset to shortlist those with stronger correlations as numerical predictors for our movie rating.
4.Build a linear regression model to analyze the relationship between the outcome and predictor variables.
5.We plan to use another model (to be decided later) to compare the accuracy between the two models.


## 2. Data
```{r}
data <- read.csv("../data/movie_metadata.csv")
glimpse(data)
```
Description of the Columns:

color :- Movie is black or coloured
director_name:- Name of the movie director
num_critic_for_reviews :- No of critics for the movie
duration:- movie duration in minutes
director_facebook_likes:-Number of likes for the Director on his Facebook Page
actor_3_facebook_likes:- No of likes for the actor 3 on his/her facebook Page
actor2_name:- name of the actor 2
actor_1_facebook_likes:- No of likes for the actor 1 on his/her facebook Page
gross:- Gross earnings of the movie in Dollars
genres:- Film categorization like ‘Animation’, ‘Comedy’, ‘Romance’, ‘Horror’, ‘Sci-Fi’, ‘Action’, ‘Family’
actor_1_name:- Name of the actor 1
movie_title: Title of the movie
num_voted_users:-No of people who voted for the movie
cast_total_facebook_likes:- Total facebook like for the cast of the movie
actor_3_name:- Name of the actor 3
facenumber_in_poster:- No of actors who featured in the movie poster
plot_keywords:-Keywords describing the movie plots
movie_imdb_link: Link to the movie on IMDB
num_user_for_reviews: Number of users who reviewed the movie
language: Language of the movie
country: Country of the movie
content_rating: Rating of the movie
budget: Budget of the movie
title_year: Year the movie was released
actor_2_facebook_likes: Number of likes for actor 2 on Facebook
imdb_score: IMDB score of the movie
aspect_ratio: The ratio of width and height of the screen
Movie_facebook_likes: Number of likes for the movie on Facebook


## 3. Data analysis plan

Q1: The outcome (response, Y) and predictor (explanatory, X) variables you will use to answer your question

Ans. 
The outcome variable is the rating of the movie and the predictor variables are budget,director,actor and genre.

Q2: The comparison groups you will use, if applicable

Ans.
We are not using comparison groups yet

Q3: Very preliminary exploratory data analysis, including some summary statistics and visualizations, along with some explanation on how they help you learn more about your data. (You can add to these later as you work on your project.)

For the preliminary exploratory data analysis, we will identify the NaN values by using is.na() and use head() function to display the first few roles for the data cleansing. After that we will use the summary statistics and the visualization to find the patterns and outliers in the dataset. We will use function summary() as the approach to find the summary statistics which includes the minimum, maximum, mean, and standard deviation. For the visualization, we will use the ggplot library to create the scatter plot between the title year and the IMDB score by using the color to indicate each color of the movie and the bar plot between the title year and the average budget that grouped by the title year which helps us find the patterns of the dataset. 

```{r}
#Returning the first few parts of the dataframe
head(data)
```
```{r}
#Using summary() function to get the statistics of the dataframe
summary(data)
```
```{r}
#Finding the variables which have missing values (NA's) in the dataframe so that we could eliminate or modify the missing values later as we work on exploring our 4 main predictor variables.
colSums(is.na(data))
```

A ggplot object is initialized by ggplot(). It can also be used to indicate the set of plot aesthetics meant to be shared by all succeeding layers unless specifically changed, as well as to declare the input data frame for a graphic.

The visualization below is describing IMDb score over a century for Black and white movie versus color pictures. Color picture has much higher ratings and the trend is clearly visible below through the scatter plot.
We wanted to see if Black&White movies still hold any significance with the progression of movies.
```{r}
  ggplot(data = data, aes(x = title_year, y = imdb_score, color = color)) +
  geom_point()
```
```{r}
#We find other than Color and Black&White movies we have a blank data in the unique value for color of movies. We have not transformed it as we aren't convinced that color would play an important role in rating prediction for a movie.
unique(data$color)
```


The visualization below is describing increasing budget for various movie for over an century. With commercialization of movies, marketing costs, CGI and special effects and cast payments have increased rapidly over years and hence we can see a gradual increase in the budget.

Average Budget By Year
```{r}
ggplot(data = data, aes(x = title_year, y = budget, group = title_year)) +
  geom_bar(stat = 'summary', fun = 'mean')+
  scale_y_continuous(labels = scales::comma)

  
```

q.4 The method(s) that you believe will be useful in answering your question(s). (You can update these later as you work on your project.)
Ans. 

For the Exploratory data analysis we are planning to summary() for inspecting the data in the Dataframe, dim() function for dimension of the dataset, various charts from ggplots library for data visualization and is.na() for finding missing values in various columns.

The method which we believe would be useful for answering our question i.e using linear regression for prediction of our outcome. Here, we are planning to predict the Rating variable, which is based on several other variables like Director, Genre, Actor, Budget and hence Linear regression would be helpful for answering our question. We plan to use Predict() for prediction and Plot() for a Graphical visualization

q.5 What results from these specific statistical methods are needed to support your hypothesized answer?

Ans. We would utilize P values and coefficients in our regression analysis to demonstrate the statistical significance of our model's associations. The linear regression coefficients give a mathematical explanation of how each independent variable and dependent variable are related. 
The statistical significance of these connections is indicated by the coefficients' p values. We further plan to use Cross validation, getting new datasets for analysis etc. for supporting our answer. We would further investigate and employ additional strategies for supporting our answer.


